### 2015-05-04 related.tex
Updates from ShareLaTeX

\section{Related Work}
The connection of HOW to model-based evolutionary programming was discussed above.
In summary,   model-based tools such as NSGA-III~\cite{deb14} or MOEA/D~\cite{zhang07:TEC}
use some domain model as part of extensive Monte Carlo studies (augmented with some
evolution of selected best instances).
The premise of HOW is that we have access to domain data, but not a domain model.
Also, the results shown above were obtained without evolving/updating some current population
multiple times.

Planning with HERE is somewhat different to the recommendor systems.
As described by Robillard and Walker~\cite{rob14}, recommendor systems
in software engineering
are less {\em directive} and more    {\em collaborative} than policy
generators:
\bi
\item
Directive systems such as HERE tell developers exactly what to do
(e.g. ``use these settings in a Makefile'') 
\item
The collaborative systems discussed by Robillard and Walker
are more suggestive guides that
help (e.g.) programmers focus on very small sets
with  very large libraries of code of documentation.
\ei
One difference between directive and collaborative systems is that
the latter may not presume to suggest exactly what to do next.
However, planning systems like HOW are much more ardent: then make precise recommendations
of how to change all items in the test suite.

HOW is more a data mining method than a model-based optimizer. It represents our next attempt
to develop model-less instance-based optimizers.  Our previous attempts in this area
were the W2, WHERE, and IDEA systems~~\cite{menzies13:brady,Menzies2013:local,me12c}.
W2 was developed with Brady et al.~\cite{menzies13:brady}. 
W2 reasoned over deltas in the the raw
attributes values (and not the synthesized attributes used by HOW). 
The results of  adjusting values via those deltas was assessed via an SQL-like
query that selected all examples in the database that satisfied the newly adjusted
values. W2 suffers from two problems that are resolved
in this paper: {\em optimization failure} and {\em instance exhastion}.
\bi
\item 
As reported in that paper, W2
had a large number of {\em optimization failures}; i.e. after applying its recommendations,
the objective scores actually got worse. We recommend HOW's approach (of reasoning
over the synthesized attributes) to W2 since {\em in none of the above
results} do we observe that {\em after} we treat the data, that the objective scores are worse.
\item
As to {\em instance exhaustion}, W2 was a simple case-based reasoning system that never generalized
over its instances. Hence, the SQl-like query used to asses W2's plans often returned
a vanishingly small number of instances. To avoid this problem, we strongly recommend
combining a planning system with a prediction system (as done in this paper) such that we can
generalize to adjusted examples not seen on the original sample.
\ei
The WHERE tool~\cite{Menzies2013:local} developed the recursive clustering method over synthesized attributes
  used in this paper. WHERE was   a prediction
system since reported current patterns in the data, rather than proposing plans
on how to improve those patterns. Nor did that tool offer guidance on how to assess
the results of stepping along those patterns to guide changes to the system.

The problems of W2 and the success of WHERE suggested in might be useful to combine
WHERE's synthesized attributes with W2. The result was the IDEA tool~\cite{me12c}. 
While the initial prototype seemed promising~\cite{me12c}, we made the mistake of using
the tree structure of the recursive clusters to generate plans (in IDEA, the plan to
move an instance from one poorly performing cluster $C_1$ to another $C_2$ was computed
from the different in the  median
values of the instances in the two  branches that lead to $C_1$ and $C_2$).  When
that approach was applied to the data used in this paper, the results were quite weak
(very little net improvement). However, we abandoned the tree-query approach
and went for the pure instance-based approach discussed at the start of this paper.

Moving on to other releated work, depending on the audience  we would
 call HOW a {\em spectral learner}, 
a {\em response surface method}, or  a {\em local search operator}.
According to
Kamvar et al~\cite{kamvar03}
{\em spectral learners}reason across $m$ underlying dimensions synthesized
using (say) PCA~\cite{pearson1901}. For example,
algorithms like PDDP~\cite{boley98} use PCA to
recursively divide data into smaller regions where each level of recursion needs an   $O(N^2)$ analysis
to infer the principle components. HOW performs an analogous procedure using a   $O(2N)$ analysis
based on a heuristic proposed by
Faloutsos and Lin~\cite{Faloutsos1995} (see above, the ``FASTMAP'' algorithm).

{\em Response surface methods}  build a fast approximation to the function being optimized,
then uses that  approximation to take intelligent guesses about useful mutations. HOW
is a non-parametric response surface method that finds its response surface via recursive
clustering. For an example of a parametric response surface method, that uses Gaussian process
models, see Zuluaga et al.~\cite{zuluaga13}.

A {\em local search operator} that a some probability $p$, nudges  instance variables
along a dimension that improves its objective scores. HOW's recursive clustering methods
combine $n$ dimensions into $m \ll n$ so its ``nudges'' are simultaneous
adjustments along multiple dimensions. For an example of other kinds of local search algorithms, that only adjust
one variable at a time, see the literature on WALKSAT~\cite{Selman1992} and MAXWALKSAT~\cite{kautz97}.
See also the literature on local search methods in multi-objective optimization.
For example,
Peng et al.~\cite{peng09:ls} have augmented MOEAs with
local search  (i.e. applying a problem-specific repair/improvement
heuristic on some current solution).
Also, Igel et al.'s~\cite{igel07} multi-objective
covariance matrix adaptation evolution strategy
can run the mutations along  ``ridges'' in the search space.

Note that the above response surface methods and local search operators cannot be compared to HOW
since, unlike HOW, they assume the presence of some model that can evaluate generations of newly mutated
instances. Recall from the above that the goal of HOW is to offer an optimization method without
using the evolutionary methods that make extensive demands on an underlying domain model.