### 2015-05-06 code.tex
Updates from ShareLaTeX

\begin{figure} 
\begin{python}[right]
def HERE(trainingData, testingData):
  training(trainingData)
  return [nearest(z).displace(z) 
          for z in testing]

def training(trainingData):
  Line.using = what2Use(trainingData)
  clusters = cluster(trainingData):
  for cluster in clusters:
    one = examplar(cluster)
    two = examplar(nearest(one, clusters))
    Line(one,two)

def what2Use(trainingData):
  second  = lambda lst: lst[1]
  weights = map(second,
                sorted(featureWeighting(trainingData)))
  use    = int(len(weights)*Line.B)
  return [i for i,_ in enumerate(weights[:use])]

def nearestLine(z):
  out,least=None,10000
  for line in Line.all:
    d = line.dist(z)
    if d < least: out,least = line,d
  return out
\end{python}
\caption{Instance-based Planning using HERE.
Testing data is {\em displaced}
towards   the $X$ point of the nearest {\tt Line} (defined in \fig{line})
The {\tt geometry} method comes from \fig{code1c}.
The {\tt featureWeighting} function comes from \fig{code1a}.
The {\tt cluster} function comes from \fig{fastmapCode}.
}\label{fig:HERE}   
\end{figure}

\begin{figure} 
\begin{python}[right]
class Line:
  "Lines connect a worse cluster Y to a better one X"
  all   = []
  using = []
  F     = 0.5
  B     = 0.5
   def __init__(i,x,y):
    if x.score < y.score:
      x,y = y,x
    i.c = dist(x,y) 
    i.obj = len(x) 
    Line.all += [i]
  def displace(i,z):
    for j,(xj,zj) in enumerate(zip(x,z)):
      if not j == i.obj and  j in Line.using:
         z[j] = zj + Line.F*( xj. - zj
    return z
  def dist(i,z):
    _,out = geometry(i.x,i.y,z)
    return out
  def exemplar(i,cluster):
    #Apply the rules at end of section II.B
\end{python}
\caption{The primary data structure of HOW
is the {\tt Line}.
Training data is converted into a set of {\tt Line}s,
each of which is defined by a $X,Y$ pair, where $X$
has better scores than $Y$. }\label{fig:line}
\end{figure}

\begin{figure} 
\begin{python}[left]
def dist(x,y):
   "See Aha et al. reference [22]. XXX"

def geometry(x, y, c, z): 
  a = dist(z,x)
  b = dist(z,y) 
  x= $(a^2 + c^2 - b^2)/(2ac)$ 
  y=  $\sqrt{a^2 - max(x,0)^2)}$
  return x,y

def nearest(x,data):
  return furthest(x,data,best=1000,better=lt)

def furthest(x,data, best= -1, better=gt):  
  out = None
  for one in data:
    d = dist(one,x)
    if gt(d,best): out, best = y, d
  return out
\end{python}
\caption{Some routines for 3 point geometry,  from \fig{where}.c.}\label{fig:code1c}
\end{figure}

\begin{figure}
\begin{python}[left]
def featureWeighting(cols):
   obj = len(cols[0])
   return sorted(weight1(col,i,obj) 
                for i,col in enumerate(cols)
                if not i == obj)

def weight1(i,obj,col):
  pairs = sorted([row[i],row[obj] for row in col])
  n     = length(col)
  return sum(v*n1/n for v,n1 in recurse(pairs,[])),i

def recurse(this,cuts):
  cut,sd = divide(this)
  if cut:
    recurse(this[:cut],cuts)
    recurse(this[cut:],cuts)
  else:
    cuts += [(sd,len(this)]
  return cuts

def divide(this,tiny=2):
  lhs, rhs = Stats(), Stats(x[1] for x in this)
  n, least, cut = rhs.n*1.0, rhs.sd(), None
  for j,x in enumerate(this):
     if lhs.n > tiny and rhs.n > tiny:
       tmp = lhs.n/n*lhs.sd() + rhs.n/n * rhs.sd()
       if tmp < least:
          cut,least = j,tmp
     rhs - x[1]
     lhs + x[1]
  return cut,least

class Stats():  
    def __init__(i,inits=[]):
      i.n = i.mu = i.m2 = 0.0
      map(i.__add__,inits)
    def sd(i) :  
      return (max(0.0,i.m2)/(i.n - 1))**0.5
    def __add__(i,x):
      i.n  += 1
      delta = x - i.mu
      i.mu += delta/(1.0*i.n)
      i.m2 += delta*(x - i.mu) 
    def __sub__(i,x):
      i.n  -= 1
      delta = x - i.mu
      i.mu -= delta/(1.0*i.n)
      i.m2 -= delta*(x - i.mu) 
\end{python}
\caption{Feature weighting, from \fig{where}.d.}\label{fig:code1a}
\end{figure}

\begin{figure} 
\begin{python}[left]
def cluster(data, n,lvl=100):
  if lvl < 1: 
     return data # stop if out of levels
  leafs = [] # Empty Set
  x,y,left,right = splitAcross2Points(data) 
  if len(left) > $\sqrt{n}$:  
     leafs += cluster(left, n, lvl - 1)  
  if len(right) > $\sqrt{n}$:  
     leafs += cluster(right,n,  lvl - 1) 
  data.has = dict(x=x,y=y,
                 l=left,r=right,
                 leafs=leafs)
  return leafs

def splitAcross2Points(data): 
  tmp = random.choose(data)
  x = furthest(tmp, data)
  y = furthest(x, data) 
  c = dist(x,y)  
  if x.scores < y.scores:
     x,y = y,x 
  for one in data.members: 
    one.pos = geometry(x,y,c,one)
  data = sorted(data) # sorted by 'pos.x'
  return x, y, split(data)

def split(data):   
  mid = len(data)/2; 
  return data[mid:], data[:mid]
\end{python}
\caption{Clustering. Lines 16,17,18 come from~\cite{fastmap}. }\label{fig:fastmapCode}   
\end{figure}